Astr 513: Final Project

### 2016-05-06 Final_Project/ms.tex
Changed little things, added a bit about cosmosis

\newcommand\Planck{{\it Planck}\ }
\newcommand\Cosmosis{{\it Cosmosis}\ }

@@

\section{Introduction}
\label{sec:Intro}
Type \rom{1}a supernovae (\snea{}) have been instrumental in the discovery 
of the accelerating expansion of the universe. Presently, {\sn} surveys continue 
to be pivotal in constraining cosmological parameters. In this paper, we present 
our analysis of comparing and combining {\sna} inference of cosmological parameters 
to those from other cosmological probes. In particular, we utilize the constraints 
from the Cosmic Microwave Background (CMB) and the Baryonic Acoustic Oscillations 
(BAO) that are gathered in \citet{planck2013}. We look at two cosmologies in our 
analysis: the $\Lambda \textnormal{CDM}$ universe and the wCDM universe. This 
paper is outlined as follows. In Section 2, the process and importance of {\snea} 
are summarized. A short description of the Planck mission is given in Section 3. 
Section 4 describe the two {\sna} surveys whose data we use for our analysis. The 
significance of using distance modulus data to constrain cosmological parameters 
is illustrated in Section 5. The CMB and BAO priors taken from \citet{planck2013} 
are also specified in Section 5. Our results and analysis on the benefit of omitting 
or including Bayesian priors from other cosmological probes 
are in Section 6 and Section 7.

@@

\citet{planck2013} report marginalized
posterior distributions for all parameters,
which we report in Table~XXX. This prior is implemented
in the planck2015 module implemented in \Cosmosis \citep{cosmosis}.

@@

Two cosmologies are of interest to us: the $\Lambda \textnormal{CDM}$ 
universe and the wCDM universe. In the $\Lambda \textnormal{CDM}$ 
universe, $w=-1$ and constraints are placed on $\Omega_M$ and $h$. 
The wCDM universe is a one-parameter extension of $\Lambda \textnormal{CDM}$ 
in which $w$ is an arbitrary parameter for a constant dark energy 
equation of state. In wCDM, constraints are placed on $\Omega_M$, $w$, 
and $h$. In both cosmologies, the universe is spatially flat 
($\Omega_k = 0$) and thus $\Omega_{\Lambda}=1-\Omega_M$.

\subsection{Likelihood}
\label{sec:likelihood}
Using the above model for the distance modulus allows us to write a
probability that a model can generate our data:
%
\begin{eqnarray}
  \label{eq:likelihood}
  P(\mu(z)|\mathrm{cosmo}) = \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\exp
  \left[-\frac{1}{2} \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
    \mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
\end{eqnarray}
%
where $\mathbf{C}$ is the full covariance matrix between all of the $N_{samples}$ 
in the data, and $\mathrm{cosmo}$ represents the cosmology that we are investigating
(e.g. flat $\Lambda$CDM, $wCDM$, etc.). 
This is not the full picture though, since we must employ
Bayes' Theorum in order to include any priors on our choice of 
parameters and recover the posterior distribution function of our
parameters given our data
%
\begin{equation}
  \label{eq:posterior}
  P(\mathrm{cosmo}|\mu(z)) = P_{prior}(\mathrm{cosmo}) P(\mu(z)|\mathrm{cosmo}).
\end{equation}
%
For this work we investigate a combination of priors. \Cosmosis is implemented
in a way that requires the user to provide bounds on the parameter space, which we
detail for each parameter in \autoref{tab:priors}
Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
\Planck priors.

>>>>

\newcommand\Planck{{\it Planck}\ }
\newcommand\CosmoSIS{{\it CosmoSIS}\ }

@@

\section{Introduction}
\label{sec:Intro}
Type \rom{1}a supernovae (\snea{}) have been instrumental in the discovery 
of the accelerating expansion of the universe. Presently, {\sn} surveys continue 
to be pivotal in constraining cosmological parameters. In this paper, we present 
our analysis of comparing and combining {\sna} inference of cosmological parameters 
to those from other cosmological probes. In particular, we utilize the constraints 
from the Cosmic Microwave Background (CMB) and the Baryonic Acoustic Oscillations 
(BAO) that are gathered in \citet{planck2013}. We look at two cosmologies in our 
analysis: the $\Lambda \textnormal{CDM}$ universe and the wCDM universe. This 
paper is outlined as follows. In Section 2, the process and importance of {\snea} 
are summarized. A short description of the \Planck mission is given in Section 3. 
Section 4 describe the two {\sna} surveys whose data we use for our analysis. The 
significance of using distance modulus data to constrain cosmological parameters 
is illustrated in Section 5. The CMB and BAO priors taken from \citet{planck2013} 
are also specified in Section 5. Our results and analysis on the benefit of omitting 
or including Bayesian priors from other cosmological probes 
are in Section 6 and Section 7.

@@

\citet{planck2013} report marginalized
posterior distributions for all parameters,
which we report in Table~XXX. This prior is implemented
in the planck2015 module implemented in \CosmoSIS \citep{cosmosis}.

@@

Two cosmologies are of interest to us: the  $\Lambda$CDM 
universe and the $w$CDM universe. In the  $\Lambda$CDM 
universe, $w=-1$ and constraints are placed on $\Omega_M$ and $h$. 
The $w$CDM universe is a one-parameter extension of  $\Lambda$CDM 
in which $w$ is an arbitrary parameter for a constant dark energy 
equation of state. In $w$CDM, constraints are placed on $\Omega_M$, $w$, 
and $h$. In both cosmologies, the universe is spatially flat 
($\Omega_k = 0$) and thus $\Omega_{\Lambda}=1-\Omega_M$.

\subsection{CosmoSIS}
\label{sec:CosmoSIS}
\LC{I'll write more later if need be.}
\CosmoSIS is a framework that connects multiple existing codes as modules in order to do cosmological parameter estimation. Users can utilize existing codes or use their own codes for modules. For our analysis, \CosmoSIS provides a convenient way to link together the Code for Anisotropies in the Microwave Background (CAMB), the \Planck likelihood code (our priors), and our {\sn} likelihood code in a pipeline.

\subsection{Likelihood}
\label{sec:likelihood}
Using the above model for the distance modulus allows us to write a
probability that a model can generate our data:
%
\begin{eqnarray}
  \label{eq:likelihood}
  P(\mu(z)|\mathrm{cosmo}) = \frac{1}{\sqrt{2\pi|\mathbf{C}|}}\exp
  \left[-\frac{1}{2} \sum_{i,j=0}^{N_{samples}}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_i 
    \mathbf{C}^{-1}_{i,j}\left(\mu(z) - \mu_{\mathrm{model}}(z)\right)_j\right]
\end{eqnarray}
%
where $\mathbf{C}$ is the full covariance matrix between all of the $N_{samples}$ 
in the data, and $\mathrm{cosmo}$ represents the cosmology that we are investigating
(e.g. flat $\Lambda$CDM, $w$CDM, etc.). 
This is not the full picture though, since we must employ
Bayes' Theorum in order to include any priors on our choice of 
parameters and recover the posterior distribution function of our
parameters given our data
%
\begin{equation}
  \label{eq:posterior}
  P(\mathrm{cosmo}|\mu(z)) = P_{prior}(\mathrm{cosmo}) P(\mu(z)|\mathrm{cosmo}).
\end{equation}
%
For this work we investigate a combination of priors. \CosmoSIS is implemented
in a way that requires the user to provide bounds on the parameter space, which we
detail for each parameter in \autoref{tab:priors}
Within those bounds, we investigated using a flat prior $P_{prior}=1$ and
\Planck priors.
